On Supporting Molecular Visualization on Grid Resources using Visual Programming Environments

### 2008-04-24 opal/documents/grid2008/paper.tex
added conclusions. now only missing the rest of the paper :)

- Opal wrapper and GUI \cite{opal}: Sriram

- Opal and Opal-CSF4: Sriram

- Data services: Sriram

@@

@@

- Conclusions (Sriram)

>>>>

\subsection {Opal Overview}

- Opal wrapper and GUI \cite{opal}: Sriram 
- Opal and Opal-CSF4: Sriram
- Data services: Sriram

\subsection {Towards Opal2}

@@

- PDB2PQR-APBS workflow in action

@@

Visual programming environments enable non-programmers to intuitively and
interactively build scientific workflows and visualize their data, without
having to write code or understand the details of the back-end
infrastructure. In this paper, we described our tool called Vision, which
is a component-based, application-domain agnostic, cross-platform visual
programming environment. On the other hand, Grid computing provides an
enormous opportunity to scientific end-users to use distributed computing
and storage resources to enable novel science hitherto impossible using
traditional computing mechanisms. However, most Grid middleware tools are
complicated to install and too jargon rich for scientists to learn. In
order to simplify the use of Grid infrastructures, one has to provide easy
to use abstractions to the complex back-end functionality. Towards this
end, we have developed the Opal toolkit, which enables wrapping scientific
applications running on Grid resources as simple Web services, which can be
accessed via a number of user interfaces. In this paper, we described how
we can use the metadata provided by the Opal services to automatically
generate Vision nodes to access the applications running on Grid resources.
With the help of a molecular visualization workflow, we demonstrated how
this is being used by domain scientists, while being completely shielded
from the complexity of the back-end computational infrastructure. We
believe that the use of services-oriented science within visual programming
environments will go a long way towards lowering the entrance barrier for
scientific end-users to effectively use Grid middleware.

The authors would like to acknowledge support from the NIH P41 RR 08605 to
NBCR; TATRC W81XWH-07-2-0014 to PWA, the Gordon and Betty Moore Foundation
grant for the CAMERA project, and the NSF grants INT-0314015 and
OCI-0627026 for the PRAGMA project.
